Exact Symbolic Dynamic Programming for Continuous POMDPs

### 2012-05-22 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
new 10 page draft

\begin{abstract}
Decision-theoretic planning problems try to find the optimal sequence of actions for a given environment. Using continuous models in such problems is naturally closer to the real-world. Partially Observable MDPs model uncertain environments where the underlying state is not be fully observable.

While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations. This symbolic dynamic programming approach avoids enumerating the state and observation space providing solutions for our general discrete and continuous partially observable MDP. The main contribution of this paper is to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions associated with each state partition. This allows us to perform the Bellman backup operation efficiently. We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
\end{abstract}

\section{Introduction}
% Write intro here
In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them. Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in abstract problems. In many real-world problems, resources and sensor inputs are naturally modeled using continuous variables.

Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain. Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}. Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem. In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment. This can quickly become infeasible even for small problem sizes.

@@

\textbf{Example} \textsc{(Power Plant)}
\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity. Feed-water is exposed to heat in the tubes leading to the water tank . Evaporating the water is performed under specific pressure and temperature inside the attached water tubes. 
Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain. For this reason, the process is modeled using POMDP in the literature. The temperature is assumed to be a continuous variable that is not observed. Noisy observations of the temperature are used to decide on the action of closing or opening the valve.}

@@

Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state%$P(\vec{d}',\vec{x}'|\cdots,a)$
; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.  A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.

%%%% forms of transition, observation and reward function
To define the transition function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN). Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition to be independent:

{\footnotesize
\begin{equation}
P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
\end{equation}}
The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
{\footnotesize
\begin{equation}
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{equation}}

The \emph{binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. For \emph{continuous} variables
the CPF $P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ is represented by \emph{piecewise
linear equations} (PLEs) that are first-order Markov and deterministic, i.e. the probabilities are encoded using the Dirac $\delta[\cdot]$ function.  The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency via bilinear solvers.

For the \textsc{Power Plant} example, the temperature $t_s \in \mathbb{R}$ is modelled as the continuous state variable. Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$. 
For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=open) &: t_s - 5 \\ 
(a \neq open) &: t_s + 7 \\
\end{cases}
\right]\nonumber
, 
R(t_s,a) = 
\begin{cases}
 (a=open) &: -1 \\
(a \neq open)\wedge (t_s>T) &: -1000 \\
(a \neq open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}\nonumber
\end{align}
%contribution? 
This example will try to maximize its reward using observations on the temperature. The observations recieved before closing the plant to a safe temperature are therefore critical.  For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following:  
\begin{align}
P(t_{o_1}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.9 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
\end{cases}\nonumber
, 
P(t_{o_2}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.1 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
\end{cases}\nonumber
\\
P(t_{o_1}|t_s',close) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.1 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
\end{cases}\nonumber
,
P(t_{o_2}|t_s',close) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.9 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
\end{cases} \nonumber
\end{align}

After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains. For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
\begin{align}
P(t_o|t_s',open) = 
\begin{cases}
 (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
 \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
\end{cases}\nonumber
\end{align}

@@

\section{DC-POMDP solution}
The DC-POMDP will choose actions in order to compute an optimal plan over time. This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state. In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$. Observations don't provide a unique state therefore the probability of the state is stored as the belief state $b$.

 At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$  where for simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$).
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
%The next section provides the mathematics required to solve this equation in DC-POMDPs.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
\end{center}
\vspace{-2mm}
\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The optimal value function for \textsc{Power Plant}
as a decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
%\end{center}
%\vspace{-3mm}
%\caption{%\footnotesize 
%The optimal value function for \textsc{Power Plant}
%as a decision diagram: 
%the \emph{true} branch is solid, the \emph{false}
%branch is dashed.} 
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Dynamic Programming solution} 
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
The decision at each time step is to pick an action based on past actions and observations. The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1. %From here on we assume this finite set of observation partitioning as $\mathcal{O}$.

We define the value iteration algorithm for discrete observations and discrete states that finds the optimal value for each horizon starting with an initial belief. Although the number of belief states is infinite, the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$. At each of the $h$-steps the value function is parameterized by the set of  $\alpha$-vectors that partition the belief state. 
At a particular belief point, we can compute the next value function: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + \gamma \sum_o p(xd_o|a,b)V^h(b|xd_o,a)\right)\nonumber
\end{align*}
Replacing the previous iteration value with (3) and also updating the belief after executing action $a$ and observing $xd_o$ results in the following equation: 
\begin{align*}
V^{h+1}(b) = \max_a \left( b \cdot r_a + 
\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right)\nonumber
\end{align*}
where $r_a = r(x_s,d_s,a)$ and  
$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$

Calculating all possible ways of building the value function independent of a certain belief state can define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. 
Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (4):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
\begin{align}
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
For each horizon, the optimal $\alpha$-vector is obtained from the maximum over all $\alpha$-vectors: 
$\alpha^* = \argmax_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot b $

Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors. Thus the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation. For this reason even for a small number of discrete observations, tractable solutions may be hard to define. Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where continuous observations and states are allowed in the DC-POMDP. We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree. We present the symbolic value iteration algorithm in the case of continuous observations using relevant observation partitioning of this continuous space.

@@

\section{Conclusion}
We presented the first exact solution to continuous states and observations in a DC-POMDP framework. Transitions and reward functions are defined as piecewise linear functions of the continuous states. Continuous observations require definition of explicit partitions based on the observation model and the belief state of the agent. The key contribution is to define the partitions on the observation space and derive the related probabilities using symbolic integration and maximization. We extend the application to more than one state-observation setting and present the results.

>>>>

\begin{abstract}
Decision-theoretic planning searches for the optimal sequence of actions for a given uncertain environment. Partially Observable MDPs (POMDP) can model such environments where the underlying state is not be fully observable. %While discrete models of the environment is often used for planning, continuous models are naturally closer to the real-world. 
While previous work have provided solutions for discrete states and/or observations, here we define the first exact solution to both continuous state and observations. The symbolic dynamic programming approach in this paper avoids enumerating the state and observation space providing solutions for general multi-variate discrete and continuous POMDPs. Our main contribution is that for a fixed set of reachable belief states and a fixed set of actions, only a \textit{finite} partitioning of the continuous observation space is required for optimality. 
Using function representation of Affine algebraic decision diagrams (XADDs) and symbolic case operators we derive the automatic partitioning for piecewise linear dynamics %such as the reward, transition and observation function. 
We demonstrate empirical results on an operation research domain to show the \textit{first exact solution} to continuous state and observation POMDPs.  
%to use continuous states with discrete observations and extending it to continuous observations by defining relevant observation partitions %associated with each state partition. %This allows us to perform the Bellman backup operation efficiently. 
%We address the problems in POMDP continuous planning and solutions to solve them symbolically. %To compactly represent the exact value functions Affine algebraic decision diagrams are used. 
\end{abstract}

\section{Introduction}
% Write intro here
In many complex systems complete information of the state is not available due to noisy or poor sensors and readings of them. Partially observable Markov Decision Processes (POMDPs)\cite{POMDPfirst} can be used for planning under such uncertainty in such problems. The solution to a POMDP is to find an answer to the fully-observable belief-MDP, whose states are probability distributions over the state space of the original POMDP \cite{kaebling}
Real-world resources and sensor inputs are often naturally modelled using continuous variables. 
Unfortunately most POMDP techniques can not be applied to such settings without assuming prior knowledge about the problem domain. Most methods use some approximation to discretize the continuous state, observation or action space which increases computational complexity \cite{Thrun99h}. The optimal value function of belief-MDPs is piecewiselinear and convex in these discrete cases \cite{sondik71} but has also been proven in the setting of only continuous states. \cite{PerseusObs}

Exact solutions to a multi-variate continuous state and observation setting has not been approached for the curse of dimensionality problem. In assuming a continuous observation setting, the agent must search for all possible observations that are read from the environment. This can quickly become infeasible even for small problem sizes.

@@

\textbf{Example} \textsc{(Power Plant)}
\emph{The steam generation system of a power plant aims to provide hot steam to the steam turbine which in turn produces electricity. Feed-water is exposed to heat in the tubes leading to the water tank . Evaporating the water is performed under specific pressure and temperature inside the attached water tubes. 
Mixing water and steam pressures makes reading tank levels and temperatures very hard and uncertain. For this reason, the process is modeled using POMDP in the literature. The temperature $t_s \in \mathbb{R}$ is modelled as a continuous state variable. Noisy observation of the temperature $t_o \in \mathbb{R}$ is used to help decide the optimal action to either open or close the pressure valve $a \in \{open,close\}$.}

@@

Three functions are required for modeling DC-POMDPs: (1) $\mathcal{T}: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow  [ 0, 1 ]$ a Markovian transition model defined as the probability of the next state given the action and previous state %$P(\vec{d}',\vec{x}'|\cdots,a)$
; (2)  $\mathcal{R}:\mathcal{S}\times\mathcal{A} \rightarrow \mathbb{R}$ a reward function which returns the immediate reward of taking an action in some state and (3) an observation function defined as $\mathcal{Z} : \mathcal{S} \times \mathcal{A} \times \mathcal{O} \rightarrow [ 0, 1 ]$  which gives the probability of an observation given the outcome of a state after executing an action.  A discount factor $\gamma, \; 0 \leq \gamma \leq 1$ is included in the model to account for future rewards in higher horizons.

%%%% forms of transition, observation and reward function
To define the transition and observation function, we use the fact that the state variables of this POMDP can be factored into a Dynamic Bayesian Network (DBN). Not allowing synchronic arcs between the discrete and continuous variables defines their joint transition or observation to be independent: 
%{\footnotesize
%\begin{equation}
%P(\vec{d_s}',\vec{x_s}'|\vec{d_s},\vec{x_s},a) = 
%\prod_{i=1}^n P(d_{s_i}'|\vec{d_s},\vec{x_s},a) \prod_{j=1}^m P(x_{s_j}'|\vec{d_s},\vec{d_s}',\vec{x_s},a). \nonumber 
%\end{equation}}
%The joint observation function can also be defined considering the direct correspondence between the state and observation variables:
{\footnotesize
\begin{equation}
P(\vec{d_o},\vec{x_o}|\vec{d_s},\vec{x_s},a) = 
\prod_{i=1}^n P(d_{o_i}|d_{s_i},a) \prod_{j=1}^m P(x_{o_j}|x_{s_j},a). \nonumber 
\end{equation}}
\emph{Binary} variables are represented using conditional probability functions (CPF) in the form of $P(d_i'|\vec{d},\vec{x},a)$. \emph{Continuous} variables
$P(x_j'|\vec{d},\vec{d'},\vec{x},a)$ are represented by \emph{piecewise
linear equations} (PLEs) that are first-order Markov and deterministic, i.e. the probabilities are encoded using the Dirac $\delta[\cdot]$ function. The reward function $R(\vec{d},\vec{x},a)$ is set to be any general piecewise linear  function that could facilitate pruning for space efficiency.% via bilinear solvers. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figure 1 - policy tree
\begin{figure}[t!]
%\vspace{-1mm}
\begin{center}
\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
\includegraphics[width=0.45\textwidth]{pics/b1-4.pdf}
\end{center}
\vspace{-2mm}
\caption{\footnotesize Left: Example conditional plan $\beta^h$ for discrete observations. Right: The optimal value function for \textsc{Power Plant}
as a decision diagram: 
the \emph{true} branch is solid, the \emph{false}
branch is dashed.}
\label{fig:cond_plan}
\vspace{-1mm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[t]
%\begin{center}
%\includegraphics[width=0.4\textwidth]{pics/cond_plan2.pdf}
%\end{center}
%\vspace{-3mm}
%\caption{%\footnotesize 
%The optimal value function for \textsc{Power Plant}
%as a decision diagram: 
%the \emph{true} branch is solid, the \emph{false}
%branch is dashed.} 
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the \textsc{Power Plant} example, the transition and reward functions are defined as below:
% For the transition probability, we use the Dirac function to model the deterministic equations. We can also define stochasticity in the transition model using boolean random variables that are sampled stochastically.  The reward function can be any linear function of the state or action. Going above a threshold temperature (e.g. $T=10$) will cause an explosion in the plant when trying to close the valve which results in the negative reward of $-1000$. Staying below this temperature is safe and will produce electricity an gain the reward of $100$. The reward of an open valve is $-1$.
{\footnotesize
\begin{align}
P(t_s'|\vec{t_s},a)= \delta\left[ t_s' - 
\begin{cases}
 (a=open) &: t_s - 5 \\ 
(a \neq open) &: t_s + 7 \\
\end{cases}
\right]\nonumber
\hspace{5mm}
R(t_s,a) = 
\begin{cases}
 (a=open) &: -1 \\
(a \neq open)\wedge (t_s>T) &: -1000 \\
(a \neq open)\wedge \neg(t_s>T) &: 100 \\
\end{cases}\nonumber
\end{align}
}
%contribution? 
This example will try to maximize its reward using observations on the temperature. The observations received before closing the plant to a safe temperature are therefore critical.  For the observation model we begin with a discrete model for two observations $t_{o_1},t_{o_2}$ such as the following for the $open$ action:  
{\footnotesize
\begin{align}
P(t_{o_1}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.9 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.1 \\
\end{cases}\nonumber
\hspace{5mm}
P(t_{o_2}|t_s',open) = 
\begin{cases}
 \delta\left[ t_s <= 15 \right] &: 0.1 \\
 \delta\left[ \neg (t_s <= 15)\right] &: 0.9 \\
\end{cases}\nonumber
\end{align}
}
%\\
%P(t_{o_1}|t_s',close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.1 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.8 \\
%\end{cases}\nonumber
%,
%P(t_{o_2}|t_s',close) = 
%\begin{cases}
% \delta\left[ t_s <= 15 \right] &: 0.9 \\
% \delta\left[ \neg (t_s <= 15)\right] &: 0.2 \\
%\end{cases} \nonumber
%\end{align}
After defining the solution to the discrete observation with a continuous state DC-POMDP, we can extend this to consider continuous observations using a continuous uniform noise distribution which is more realistic in many domains. %For a certain interval, the probability is a fixed number, while it is zero elsewhere. The integral of the uniform distribution should sum to one. 
{\footnotesize
\begin{align}
P(t_o|t_s',open) = 
\begin{cases}
 (t_o>t_s-5) \wedge (t_o<t_s+5) &: 0.1 \\
 \neg((t_o>t_s-5) \wedge (t_o<t_s+5)) &: 0 \\
\end{cases}\nonumber
\end{align}
}

@@

\section{DC-POMDP solution}
The DC-POMDP will choose actions in order to compute an optimal plan over time. This plan is the policy $\pi$ of the agent  depending on the history of actions and observations over a Markovian state. In the partially observable case where there is uncertainty on the states, the policy is a function over a continuous set of probability distributions over $|\mathcal{S}|$ known as the belief state $b = P(x_s,d_s)$.

At any of the horizon steps, a DC-POMDP policy $\pi(b)$ is defined as the iterative plan on executing an action and observing all probable outcomes of the set of observations from the $(h-1)$-step. The belief state is updated after this according to the Bayes rule: $\sum_{d_s}\int_{x_s} b(xd_s)T(xd'_s,xd_s,a)Z(xd'_s,xd_o,a)$. \footnote{For simiplicity we use ($xd_s = x_s,d_s$) and ($xd_o=x_o,d_o$)}
The optimal policy is defined over the belief state as the sum of the expected discounted rewards over horizon $h$ starting with belief $b_0$.
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = E_{\pi^*} \left[ \sum_{h=0}^{H} \gamma^t \cdot r_h \Big| \vec{b}_0 = \vec{b} \right]\nonumber
\end{equation}
}
Solving the value function for the optimal policy is performed using Dynamic programming algorithms such as value iteration. 
%The decision at each time step is to pick an action based on past actions and observations. 
The possible set of conditional plans $\beta^h$ correspond to a decision tree. Figure 1 left, shows the decision tree for the simple case of three horizons with discrete observations. It represents a conditional plan consisting of an action and two possible observation strategies. For continuous observations, conditional plans are decision trees with infinitely many branches. In our solution we show that instead of a policy tree with infinite branches, we can create partitions based on relevant observation values, resulting in a policy tree similar to that of figure 1. %From here on we assume this finite set of observation partitioning as $\mathcal{O}$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\incmargin{.5em}
\linesnumbered
\begin{algorithm}[t!]
\vspace{-.5mm}
\dontprintsemicolon
\SetKwFunction{backup}{Backup}
\SetKwFunction{genObs}{GenRelObs}
\SetKwFunction{prune}{Prune}
\SetKwFunction{remapWithPrimes}{Prime}
\Begin
{
   $V^0:=0, h:=0$\;
   \While{$h < H$}
   {
       $h:=h+1, \Gamma^h :=\emptyset$\;
       \ForEach {$a \in A$}
       {
			$\Gamma_{a}^h :=\emptyset$ \;       		
       		\If {$ContObs =true$}
       			{$P(xd_{o_i}|xd_{s_i}) \,:=\,$ \genObs{$\Gamma^{h-1},a,b_i$}\;}
       		 \ForEach {$xd_{o_i} \in \textsc{O}^{h-1}$}
       		 {
				\ForEach {$\alpha_j^{h-1} \in \Gamma^{h-1}$}
       			{
   	 		  		$\alpha_j^{h'}=$ \remapWithPrimes{$\alpha_j^{h}$} \;
   	 		  		\emph{// $\forall$ $d_i \to d_i'$ and $\forall x_i \to x_i'$} \; 
   	 		    	$\Gamma_{a,xd_{o_i},j}^h \,:=\, P(xd_{o_i}|xd_{s_i}) \otimes$ \backup{$\alpha_j^{h'},a$}\;
       	      	}
       	      	$\Gamma_{a,xd_o}^h\,:=\,\textrm{\large $\boxplus$} \Gamma_{a,xd_{o_i}}^h$\;
       	     }
           $\Gamma_a^{h} \,:=\,R_a \oplus \gamma \cdot \Gamma_{a,xd_o}^h$\;
            $\Gamma^{h} \,:=\, \Gamma^{h} \cup \Gamma_a^{h}$\;
        }  
        	%monahan's pruning first generates all vectors, then prunes
              $\Gamma^h \,:=\, $\prune{$\Gamma^h$} \;
              $V^h \,:=\, \mathrm{max}_{\alpha_j \in \Gamma^h} b_i \cdot \alpha_j$\;
              $\pi^{*,h} \,:=\, \argmax_{a} \, \Gamma_a^{h}$\;

       \If{$V^h = V^{h-1}$}
           {break $\,$ \emph{// Terminate if early convergence}\;}
   }
     \Return{$(V^h,\pi^{*,h})$} \;
}
\caption{\footnotesize \texttt{VI}(DC-POMDP, $H$,$ContObs$, $b_i$) $\longrightarrow$ $(V^h,\pi^{*,h})$ \label{alg:vi}}
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For discrete observations and discrete states the number of belief states is infinite, but the optimal value function for $h$ is proven to be a piecewise linear and convex function of the belief state \cite{smallwoodSondik}. The optimal value function is the maximum value over a finite set of alpha vectors  $\alpha^h_i$:%that define the hyperplanes:
{\footnotesize
\begin{equation}
V^h_{\Pi^*}(\vec{b}) = \max_{\alpha^h_i \in \Gamma^h} \alpha^h_i \cdot \vec{b}
\end{equation}
}
where the belief states $\vec{b}$ and each $\alpha^h_i \in \Gamma^h$ are of dimension $|\mathcal{S}|$.

Point-based dynamic programming backups compute the optimal value function given a belief state $b$. The best conditional plan for the given belief, corresponds to an $\alpha$-vector which is computed using point-based backups: 
{\footnotesize
\begin{align*}
\alpha^{h+1}(b) = R(b) + \gamma \int_{xd_o} P(xd_o|a,b)\alpha^{h+1}(b)\nonumber
\end{align*}
}
where $P(xd_o|a,b) = \int_{xd_s}\int_{xd'_s}b(s)P(xd'_s|xd_s,a)P(xd_o|xd'_s,a)$.

Calculating all possible ways of building the value function independent of a certain belief state can also define the optimal value function. To derive the $\Gamma$-set of all possible $\alpha$-vectors for each horizon, we use the value iteration algorithm from ~\cite{monahan82}. Initializing  $\alpha^0_1 = \vec{0}$ and $\Gamma^0 = \{ \alpha^0_1 \}$, $\Gamma^h$ is obtained from $\Gamma^{h-1}$ using the backup operation defined in (2):\footnote{The $\textrm{\large $\boxplus$}$ of sets is defined as 
$\textrm{\large $\boxplus$}_{j \in \{ 1,\ldots, n \} } S_j = S_1 \textrm{\large $\boxplus$} \cdots \textrm{\large $\boxplus$} S_n$ where the pairwise cross-sum $P 
\textrm{\large $\boxplus$} Q = \{ \vec{p} + \vec{q} | \vec{p} \in P, \vec{q} \in Q \}$.}
{\footnotesize
\begin{align}
V^{h+1}(b) &= \max_a \left( b \cdot r_a + 
\gamma \sum_o \max_{\langle g^h_{a,o,j} \rangle_j}  b \cdot g^h_{a,o,j} \right) \\
\Gamma^{h}_a   &= r_a + \gamma \textrm{\large $\boxplus$}_{o \in \mathcal{O}} \left\{ g^h_{a,o,j}(\cdot) \right\}_j ;  \forall \alpha^{h-1}_{j} \in \Gamma^{h-1}  \\
\Gamma^h  &= \bigcup_a \Gamma^h_a 
\end{align}
}
where $r_a = r(x_s,d_s,a)$ and  
$g^h_{a,o,j} =  \int_{x_{s'}} \sum_{d_{s'}} p(xd_o|xd'_s,a)p(xd'_s|xd_s,a) \alpha^h_j(xd'_s)$
%Each action-dependent $\Gamma$ sums over the possible set of $g^h_{a,o,j}(\cdot)$ given an observation model to derive discrete observations $xd_o$ and for all $\alpha^{h-1}_{j} \in \Gamma^{h-1} $. The value $g^h_{a,o,j}(\cdot)$ is computed by summing over all states given the probability functions and previous $\alpha$-vectors.  
Lines 9--16 of the \textsc{VI} algorithm show the above in a general algorithm used for both discrete and continuous observations. 
For all $\alpha$-vectors in the current $\Gamma$-set first an LP-solver is used to prune any inconsistency in the branches of this vector in line 17. A pairwise dominance test is performed where for all $j,k$: $\alpha_j(\vec{s}) \geq \alpha_k(\vec{s})$
The result of this pairwise test is a dominant $\alpha$-vector which is added to the set of $\Gamma^h$. 
%Here the cross-sum operation takes the set of $\alpha$ -vectors for each observation and produces a sum over vectors that depends on the number of previous iteration vectors.

Note that the $\alpha$-vector set grows exponentially in the number of observations, i.e., $|\Gamma^{h}| = |\mathcal{A}||\Gamma^{h-1}|^{|\mathcal{O}|}$ during the cross-sum operation. For this reason even for a small number of discrete observations, tractable solutions may be hard to define. Here we do not restrict our observations to discrete ones, but extend it to a fully continuous approach where lines 7--8 will be executed to generate relevant observation partitions based on a belief state $b_i$. We use the number of  observation partitions $|\mathcal{O}|$ as the branching factor in the policy tree. We present the symbolic approach required to partition this continuous space.

@@

%\section{Conclusion}
We presented the first exact solution to continuous states and observations in a DC-POMDP framework. Transitions and reward functions are defined as piecewise linear functions of the continuous states. Continuous observations require definition of explicit partitions based on the observation model and the belief state of the agent. The key contribution is to define the partitions on the observation space and derive the related probabilities using symbolic integration and maximization. We extend the application to more than one state-observation setting and present the results.
